Synthesizing Symmetric Lenses

### 2018-06-07 papers/SymmetricLenses/main.tex
Comments on the intro

\begin{abstract}
\saz{This abstract too long and doesn't get to the point quickly enough. We
  should tighten it up considerably.}
  Users commonly need to synchronize the data shared between formats, so edits
  to the data in one format gets propagated to the data in the other format.
  Manually writing the functions for these data synchronizers is tedious and
  error prone. Instead of manually writing these synchronizers, bidirectional
  languages can be used to express \emph{lenses}, groups of synchronization
  functions with guarantees about round-trip behavior. While these bidirectional
  languages help make programming these synthesizers less error-prone, they
  remain difficult to program in. Synthesizing bidirectional languages allows
  users to use lenses without manually programming them. Current
  synthesizers work only on the simplest types of lenses, when the two formats
  describe exactly the same data; this is often not the case.

  Oftentimes, each format may contain information the other lacks. Such
  synchronization tasks are described by one of the richest classes of lenses,
  symmetric lenses. However, symmetric lenses in their current formulation
  require external state in the form of a \emph{complement}. This statefulness
  makes symmetric lenses both difficult to use and difficult to synthesize.

  In this work, we reformulate symmetric lenses as \emph{simple symmetric
    lenses}. While this formulation is less expressive than symmetric lenses, it
  expresses many symmetric lenses while requiring \emph{no external state}. We
  provide combinators for simple symmetric lenses on string data, and integrate
  these combinators into Boomerang while providing full backwards compatibility.

  We develop a type-directed synthesis algorithm for generating lenses in this
  language from regular expressions describing the data formats. A significant
  technical challenge in this work arises from the fact that our target language
  permits synchronization between formats that don't have exactly the same data
  -- data in one format need not be present in the other format. This gives the
  synthesis algorithm much more freedom, as data can be completely unaligned
  (edits in one format are never propagated), completely aligned (the formats
  are in bijective correspondence), or anywhere in between. Under the assumption
  that users typically want to align as much data as possible, we provide an
  information theoretic metric for how much data is left unaligned based on the
  theory of \emph{stochastic regular expressions}. Furthermore, we use this
  metric to develop a greedy synthesis algorithm that tries to minimize
  unaligned data. We experimentally demonstrate that our algorithm can
  efficiently synthesize simple symmetric lenses on a variety of benchmark
  suites consisting of synchronizing file formats from configuration files,
  application-specific data storage files, and data cleaning tasks.

@@

% begin introduction
\section{Introduction}
% state of the world
Similar data can come in varieties of formats: scheduled jobs can be stored as
cron jobs for Linux or as Scheduled Tasks for Windows; web service resources can
be provided to clients in legacy or modern formats; calendar events can be
stored in CSV files or ICS files. Oftentimes, files should be synchronized
across these formats: when I edit a task to run more often on my Linux machine,
I want this task to run more often on my Windows machine; when I update data
through a legacy API, I want those changes to be reflected when I retrieve it
through a modern API; when I create a new calendar event in my ICS file, I want
my CSV calendar files to be updated to reflect this change.

@@

\saz{This paragraph is hard to follow --- maybe introduce $\ell : S
  \Leftrightarrow T$ and use the names $S$ and $T$?} A primary difficulty is
that there are a relatively many well-typed simple symmetric lenses. If we are
trying to find a lens between $\Regex$ and $\RegexAlt$, there is now a choice:
do $\Regex$ and $\RegexAlt$ represent the same data, and should be aligned
(changes to data in the language of $\Regex$ are propagated to synchronized
strings in the language of $\RegexAlt$), or should it be left unaligned (changes
to data in $\Regex$ update synchronized strings in $\RegexAlt$). Or instead
should certain components of $\Regex$ be aligned with certain components of
$\RegexAlt$, with the other components left unaligned.

Operating under the assumption that users wish to align as much data as
possible, our algorithm tries to align as much data as the types and examples
allow. This task gets harder when there are more complex decisions involved. For
example, if there are two fields in one format, and one in the other, which field
in the first format should be aligned to the sole field of the second format? Or
would it be better for the field on the right to be aligneded to a mish-mash
of subcomponents from each field? To answer these questions, we develop a cost
metric for reasoning about which lens aligns the most information.

@@

Our model is based on \emph{stochastic regular expressions}~\cite{?}, which
represent both a language and a probability distribution over that language
simultaneously.  We develop a way to syntactically infer the expected
information content (or \emph{entropy}) of an unambiguous regular
expression. This entropy is used to calculate the expected information content
to recover a string, when given a string it is synchronized with.

@@

We develop an algorithm that synthesizes a low-cost lens, shown in
Figure~\ref{fig:high-level-algorithm}. This algorithm consists of two
communicating synthesizers, one that proposes candidate stochastic regular
expression pairs that it thinks will align well, and one that greedily searches
for a lens between those two regular expressions that minimizes cost and does
not traverse star-related equivalences. When the first algorithm is content with
the lens it has found, it returns that lens.

Our first algorithm, \RXSearch, proposes candidate stochastic regular
expressions. These candidate stochastic regular expressions are generated by
traversing \emph{star semiring} equivalences on stochastic regular expressions.
These proposed stochastic regular expression pairs have equivalent languages and
probability distributions as the original ones: we prove that the star-semiring
rewrites this synthesizer applies are sound rewrites on stochastic regular
expressions.

@@

In summary, our contributions are:
\begin{enumerate}
\item We develop \emph{simple symmetric lenses} a new formulation for
  bidirectional programs that can project data from both sides~(\S\ref{sec:overview}),
  and describe a language that expresses such lenses on string
  data~(\S\ref{sec:ssl}).
\item We develop a means to traverse stochastic regular expression equivalences,
  while preserving probability distributions and
  languages~(\S\ref{subsec:stoch-rx}).
\item We develop stochastic DNF regular expressions and DNF
  lenses~(\S\ref{subsec:greedy-synth}). We develop a sound algorithm for
  converting stochastic regular expressions to stochastic DNF regular
  expressions and a information-theoretic cost metric on DNF lenses.
\item We extended Boomerang with our simple symmetric lens combinators, and
  integrated our algorithm with the Optician synthesis system within Boomerang,
  allowing synthesis tasks as Boomerang
  expressions~(\S\ref{sec:implementation}).
\item We evaluate our algorithm on a benchmark suite consisting of synchronizing
  file formats from configuration files, application-specific data storage
  files, and data cleaning tasks (\S\ref{sec:evaluation}).
\end{enumerate}
% end introduction

>>>>

\begin{abstract}
\saz{This abstract too long and doesn't get to the point quickly enough. We
  should tighten it up considerably.}\bcp{+1.  But not just tighten ---
  the whole story doesn't feel very compelling.}
  Users commonly need to synchronize the data shared between formats, so edits
  to the data in one format gets propagated to the data in the other format.
  Manually writing the functions for these data synchronizers is tedious and
  error prone. Instead of manually writing these synchronizers, bidirectional
  languages can be used to express \emph{lenses}, groups of synchronization
  functions with guarantees about round-trip behavior. While these bidirectional
  languages help make programming these synthesizers less error-prone, they
  remain difficult to program in. Synthesizing bidirectional languages allows
  users to use lenses without manually programming them. Current
  synthesizers work only on the simplest types of lenses, when the two formats
  describe exactly the same data; this is often not the case.

  Oftentimes, each format may contain information the other lacks. Such
  synchronization tasks are described by one of the richest classes of lenses,
  symmetric lenses. However, symmetric lenses in their current formulation
  require external\bcp{why not ``internal''?} state in the form of a \emph{complement}. This statefulness
  makes symmetric lenses both difficult to use and difficult to
  synthesize. \bcp{This paragraph seems too dense.}

  In this work, we reformulate\bcp{More a restriction than a reformulation.}
  symmetric lenses as \emph{simple symmetric 
    lenses}. While this formulation is less expressive than symmetric lenses, it
  expresses many\bcp{weak} symmetric lenses while requiring \emph{no external state}. We
  provide combinators for simple symmetric lenses on string data, and integrate
  these combinators into Boomerang\bcp{undefined} while providing full backwards compatibility.

\bcp{This bit feels more detailed / technical than needed for an abstract.}
  We develop a type-directed synthesis algorithm for generating lenses in this
  language from regular expressions describing the data formats. A significant
  technical challenge in this work arises from the fact that our target language
  permits synchronization between formats that don't have exactly the same data
  -- data in one format need not be present in the other format. This gives the
  synthesis algorithm much more freedom, as data can be completely unaligned
  (edits in one format are never propagated), completely aligned (the formats
  are in bijective correspondence), or anywhere in between. Under the assumption
  that users typically want to align as much data as possible, we provide an
  information theoretic metric for how much data is left unaligned based on the
  theory of \emph{stochastic regular expressions}. Furthermore, we use this
  metric to develop a greedy synthesis algorithm that tries to minimize
  unaligned data. We experimentally demonstrate that our algorithm can
  efficiently synthesize simple symmetric lenses on a variety of benchmark
  suites consisting of synchronizing file formats from configuration files,
  application-specific data storage files, and data cleaning tasks.

@@

% begin introduction
\section{Introduction}
% state of the world
\bcp{Not very compelling:}Similar data can come in varieties of formats: scheduled jobs can be stored as
cron jobs for Linux or as Scheduled Tasks for Windows; web service resources can
be provided to clients in legacy or modern formats; calendar events can be
stored in CSV files or ICS files. Oftentimes, files should be synchronized
across these formats: when I edit a task to run more often on my Linux machine,
I want this task to run more often on my Windows machine; when I update data
through a legacy API, I want those changes to be reflected when I retrieve it
through a modern API; when I create a new calendar event in my ICS file, I want
my CSV calendar files to be updated to reflect this change.

@@

\saz{This paragraph is hard to follow --- maybe introduce $\ell : S
  \Leftrightarrow T$ and use the names $S$ and $T$?}\bcp{The names make it a
bit better, but it is all still rather abstract.  Indeed, the whole intro is
quite abstract---could it not be driven by an example?} A primary difficulty is
that there are a relatively many well-typed simple symmetric lenses. If we are
trying to find a lens between $\Regex$ and $\RegexAlt$, there is now a choice:
do $\Regex$ and $\RegexAlt$ represent the same data, and should be aligned
(changes to data in the language of $\Regex$ are propagated to synchronized
strings in the language of $\RegexAlt$), or should it be left unaligned (changes
to data in $\Regex$ update synchronized strings in $\RegexAlt$). Or instead
should certain components of $\Regex$ be aligned with certain components of
$\RegexAlt$, with the other components left unaligned.

Operating under the assumption that users wish to align as much data as
possible, our algorithm tries to align as much data as the types and examples
allow. This task gets harder when there are more complex decisions involved. For
example, if there are two fields in one format, and one in the other, which field
in the first format should be aligned to the sole field of the second format? Or
would it be better for the field on the right to be aligneded to a mish-mash
of subcomponents from each field?\bcp{Wouldn't these questions be answered
  by examples?} To answer these questions, we develop a cost
metric for reasoning about which lens aligns the most information.

@@

Our model is based on \emph{stochastic regular expressions}~\cite{?}, which
represent both a language and a probability distribution over that language
simultaneously.  We develop a way to syntactically infer the expected
information content (or \emph{entropy}) of an unambiguous regular
expression. This entropy is used to calculate the expected information content
to recover a string, when given a string it is synchronized with.\bcp{I
know we're going to get to more technical detail later, but at this point I
am confused: a string matching a RE can in general be arbitrarily long, so
what kind of sense does it make to talk about its ``expected'' information content?}

@@

\bcp{It gets rather technical and incomprehensible for the next three
  paragraphs...}

We develop an algorithm that synthesizes a low-cost lens, shown in
Figure~\ref{fig:high-level-algorithm}. This algorithm consists of two
communicating synthesizers, one that proposes candidate stochastic regular
expression pairs that it thinks will align well, and one that greedily searches
for a lens between those two regular expressions that minimizes cost and does
not traverse star-related equivalences\bcp{???}. When the first algorithm is content with
the lens it has found, it returns that lens.

\bcp{Make clear that these algorithms are based on earlier / published ones.}Our first algorithm, \RXSearch, proposes candidate stochastic regular
expressions. These candidate stochastic regular expressions are generated by
traversing \emph{star semiring} equivalences on stochastic regular expressions.
These proposed stochastic regular expression pairs have equivalent languages and
probability distributions as the original ones: we prove that the star-semiring
rewrites this synthesizer applies are sound rewrites on stochastic regular
expressions.

@@

In summary, our contributions are:
\begin{enumerate}
\item We develop \emph{simple symmetric lenses} a new formulation for
  bidirectional programs that can project data from both sides~(\S\ref{sec:overview}),
  and describe a language that expresses such lenses on string
  data~(\S\ref{sec:ssl}).  \bcp{A new language, or a variant of
    Hofmann/Pierce/Wagner?} 
\item We develop a means to traverse stochastic regular expression equivalences,
  while preserving probability distributions and
  languages~(\S\ref{subsec:stoch-rx}).
\item We develop stochastic DNF regular expressions and DNF
  lenses~(\S\ref{subsec:greedy-synth}). We develop a sound algorithm for
  converting stochastic regular expressions to stochastic DNF regular
  expressions and a information-theoretic cost metric on DNF lenses.
\item We extended Boomerang with our simple symmetric lens combinators, and
  integrated our algorithm with the Optician synthesis system within Boomerang,
  allowing synthesis tasks as Boomerang
  expressions~(\S\ref{sec:implementation}).
\item We evaluate our algorithm on a benchmark suite consisting of synchronizing
  file formats from configuration files, application-specific data storage
  files, and data cleaning tasks (\S\ref{sec:evaluation}).  \bcp{... and
    conclude what?}
\end{enumerate}
% end introduction
